District-level male medical and traditional circumcision coverage and unmet need in sub-Saharan Africa

### 2024-01-28 paper/paper.tex
Overleaf changes

In light of this evidence, the World Health Organization (WHO) and UNAIDS recommended a scale-up in male circumcision coverage for HIV prevention in fifteen countries across Central and Western Africa with a high HIV incidence and low circumcision coverage: Botswana, Eswatini, Ethiopia, Kenya, Lesotho, Malawi, Mozambique, Namibia, Rwanda, South Africa, South Sudan Tanzania, Uganda, Zambia, and Zimbabwe \cite{UNAIDSJoint, davis2018progress, WHOVoluntary2}. Ambitious targets were set by UNAIDS to reach 90\% circumcision coverage among adolescent boys and young men aged 10--29 years by 2021 in these priority countries \cite{WHOFramework}. As such, over XX\todo{Find this number and cite} million VMMCs have been conducted in these countries between 2007 and 2022 with the support of national governments as well as international organisations such as US President’s Emergency Plan for AIDS Relief (PEPFAR).

The landscape of male circumcision in sub-Saharan Africa is complex and highly heterogeneous. Despite the upscaling in medical male circumcision (MMC) being a recent phenomenon, it has been long been practised as part of traditional male initiation ceremonies in many African countries. The provision of traditional male circumcision (TMC), such as the typical age at circumcision and what it entails, can vary considerably and is influenced by community, religious, ethnic and cultural led values. There is substantial sub-national heterogeneity in the proportion and average of men receiving TMC between at a district level in South Africa. Even on a national level TMC is universally practiced in {\color{red}\bf XXX} and is barely conducted in {\color{red}\bf XXX}. TMCs conducted during male initiation ceremonies are often performed using non-medical methods non-clinical settings by a traditional practitioner with no formal medical training. The evidence that TMC provides the same HIV prevention benefits as MMC is mixed, due to TMC may not involve a complete circumcision and instead be only partial removal of the foreskin or a simple incision in the prepuce in some populations.  \todo{Cite this paragraph.}

Why is understanding this important? DMPPT2 and Cork et al. and Thomas et al. have been approaches

In this paper,...

@@

% Geofaceted plot of 2020 coverage by age
\begin{figure}[H]
    \centering
    \includegraphics[width=.9\linewidth]
    {plots/05_map_plot_mean_circ_age.png}
    \caption{\emph{(A) Estimated distinct-level age at circumcision in 2020 in 33 sub-Saharan African countries. Maps show mean medical male circumcision (MMC) and mean traditional male circumcision (TMC) age. Countries without estimates are in grey. Mean circumcision ages are binned into three distinct groups: 0-10, 11-20 and over 20. (B) Scatterplot of district-level mean traditional male circumcision age versus mean medical male circumcision age. A diagonal dashed line passes through the origin and (30, 30). Points representing each district are coloured by whether their country is a Voluntary Male Medical Circumcision (VMMC) priority country.}}
\end{figure}

>>>>

There is a need to provide comprehensive estimates of male circumcision coverage by age,  time and type across sub-Saharan Africa, however, the landscape is complex and highly heterogeneous for numerous reasons. \todo{Need to expand this and add citations}

Firstly, the World Health Organization (WHO) and UNAIDS recommended a scale-up in male circumcision coverage for HIV prevention in fifteen countries across Central and Western Africa with a high HIV incidence and low circumcision coverage: Botswana, Eswatini, Ethiopia, Kenya, Lesotho, Malawi, Mozambique, Namibia, Rwanda, South Africa, South Sudan Tanzania, Uganda, Zambia, and Zimbabwe \cite{UNAIDSJoint, davis2018progress, WHOVoluntary2}. Ambitious targets were set by UNAIDS to reach 90\% circumcision coverage among adolescent boys and young men aged 10--29 years by 2021 \cite{WHOFramework} and as a result, over XX\todo{Find this number and cite} million VMMCs have been conducted in these priority countries between 2007 and 2022 with the support of national governments as well as international organisations such as US President’s Emergency Plan for AIDS Relief (PEPFAR).

Secondly, while the upscaling in medical male circumcision (MMC) being a recent phenomenon, it has been long been practised as part of traditional male initiation ceremonies in many African countries. The provision of traditional male circumcision (TMC), such as the typical age at circumcision and what it entails, can vary considerably and is influenced by community, religious, ethnic and cultural led values. There is substantial sub-national heterogeneity in the proportion and average of men receiving TMC between at a district level in South Africa. Even on a national level TMC is universally practiced in {\color{red}\bf XXX} and is barely conducted in {\color{red}\bf XXX}. TMCs conducted during male initiation ceremonies are often performed using non-medical methods non-clinical settings by a traditional practitioner with no formal medical training. The evidence that TMC provides the same HIV prevention benefits as MMC is mixed, as TMC may not involve a complete circumcision and instead be only partial removal of the foreskin or a simple incision in the prepuce in some populations.  \todo{Cite this paragraph.}

Cork et al. provided the first set of comprehensive estimates of circumcision coverage in sub-Saharan Africa. They used survey data within a Bayesian geo-statistical model to produce annual estimates of total male circumcision coverage subnationally in sub-Saharan Africa between 2000 and 2017. Their approach focused on predicting circumcision coverage among only the aggregate age range 15–49 years, did not distinguish between circumcision type, and did not track the structured dynamics of circumcision coverage by type within cohorts.

Thomas et al. aimed to address these issues by producing a model that synthesises both survey and programmatic data to generate region-age-time-type specific probabilities and coverage of male circumcision. The model extended a competing risks time-to-event model which enabled ... However, this study only focused on South Africa....

Here, we take a similar approach to Thomas et al. in order to model estimate the probabilities and coverage of medical and traditional male circumcision by subnational area, single-year age, and time with probabilistic uncertainty. Here

@@

% Geofaceted plot of 2020 coverage by age
\begin{figure}[H]
    \centering
    \includegraphics[width=.9\linewidth]
    {plots/05_map_plot_mean_circ_age.png}
    \caption{\emph{(A) Estimated distinct-level age at circumcision in 2020 in 33 sub-Saharan African countries. Maps show mean medical male circumcision (MMC) and mean traditional male circumcision (TMC) age. Countries without estimates are in grey. Mean circumcision ages are binned into five-year age groups and over 30s. (B) Scatterplot of district-level mean traditional male circumcision age versus mean medical male circumcision age. A diagonal dashed line passes through the origin and (25, 25). Points representing each district are coloured by whether their country is a Voluntary Male Medical Circumcision (VMMC) priority country.}}
\end{figure}
